---
title: "ASSIGMENT"
author: "ENGR. SAMUEL EFFIONG"
date: "3/20/2022"
output: 
  word_document
---

```{r setup, include=FALSE}
rm(list = ls())
library(pacman) 
library(tidyr)
library(dplyr)
library(ggplot2)
library(pander)
library(corrr)
library(scales)
library(yaml)
knitr::opts_chunk$set(echo = FALSE)
```

# Solution To Assigntment
Analysis and visual representation of **plankton** in water where shellfish are farmed
DATASET was gotten from [plankton.csv](plankton.csv)

## Plankton Dataset set for the project.
```{r}
planktondata=read.csv(file.choose(),header=T)
```
#    SOLUTION 1
*  Using **Univariate statistics** to analyse the **plankton dataset**

* FOR **MEAN,MEDIAN and MODE** in `summary statistics`

```{r}
planktondata %>%
  summarize(PLANKTONDATASET ="RESULT",Pseudonitzschia.A.Sp= mean(Pseudonitzschia.A.Sp),
            Alexandrium.Sp = mean(Alexandrium.Sp),
            Robgordia.Sp=mean(Robgordia.Sp),
            Water.Temp= mean(Water.Temp)) %>% 
  pander()


planktondata %>%
  summarize(PLANKTONDATASET ="RESULT",Pseudonitzschia.A.Sp= median(Pseudonitzschia.A.Sp),
            Alexandrium.Sp = median(Alexandrium.Sp),
            Robgordia.Sp=median(Robgordia.Sp),
            Water.Temp= median(Water.Temp)) %>% 
  pander()

planktondata %>%
  summarize(PLANKTONDATASET ="RESULT",Pseudonitzschia.A.Sp= sd(Pseudonitzschia.A.Sp),
            Alexandrium.Sp = sd(Alexandrium.Sp),
            Robgordia.Sp=sd(Robgordia.Sp),
            Water.Temp= sd(Water.Temp)) %>% 
  pander()

```

#  SOLUTION 2

* A **Boxplot** to show the distribution of **Pseudonitzschia** and a second one to show 
the distribution of **water temperature.**

* **FOR PSEUDONITZSCHIA**
```{r}
planktondata %>%
  ggplot(aes(Pseudonitzschia.A.Sp)) +
  geom_boxplot() +
  labs(title = "Boxplot to show the Distribution of Pseudonitzschia ",
       x = "PLANKTONDATASET",
       y = " RESULT") +
  theme_minimal()
```

* **FOR WATER TEMPERATURE**
```{r}
planktondata %>%
  ggplot(aes( Water.Temp )) +
  geom_boxplot() +
  labs(title = "Boxplot to show the Distribution of Water Temperature ",
       x = "PLANKTONDATASET",
       y = " RESULT") +
  theme_minimal()

```

###   COMMENT

* In the **Pseudonitzschia.A.Sp** it was observe that there  were more,
*extreme outliers* than in the **water Temperature.**

*  IN the **water temperature** plot it was observe that it was a *normal plot.*

#     SOLUTION 3

* Using **Univariate statistics** to `compare data`  for **Pseudonitzschia in the year**,
2021 with its data in previous years.

```{r}
planktondata  %>%
  group_by(year) %>%
  summarize(Pseudonitzschia.A.Sp = n()) %>%
  arrange(desc(Pseudonitzschia.A.Sp)) %>%
  pander()

```

###   COMMENT
* form the **univarite data** above,There was great improvement 
in the year 2021 `compared to other years`.

#     SOLUTION 4

* Histogram difference between **skewed distribution and normal distribution**

```{r}
ggplot(planktondata, aes(x = year)) +
  geom_histogram(fill = "cornflowerblue", 
                 color = "white", 
                 bins = 13) + 
  labs(title="histograms with a skewed distribution ", 
       y="Pseudonitzschia",
       x = "years")



ggplot(planktondata, aes(x = year)) +
  geom_histogram(fill = "cornflowerblue", 
                 color = "white", 
                 bins = 13,
                 binwidth = 8) + 
  labs(title="histograms an attribute with a normal
distribution", 
       x = "year",
       y="Pseudonitzschia")


```

###      COMMENT

*Skewed distribution was not evenly distributed.hence we can't get a symmetrical curve.

*Normal distribution was evenly distributed.hence we get a symmetrical curve.


#       SOLUTION 5
* A `bar plot` for **Species**

```{r}

plotdata <-planktondata %>%
  count(Species) %>%
  mutate(pct = n / sum(n),
         pctlabel = paste0(round(pct*100), "%"))

# plot the bars as percentages, 
# in decending order with bar labels
ggplot(plotdata, 
       aes(x = reorder(Species, -pct),
           y = pct)) + 
  geom_bar(stat = "identity", 
           fill = "cornflowerblue", 
           color = "black") +
  geom_text(aes(label = pctlabel), 
            vjust = -0.25) +
  scale_y_continuous(labels = percent) +
  labs(x = "Species", 
       y = "Percent", 
       title  = "A bar plot for Species")
```

###    COMMENT
* I choose to use Species because it has different category.
* There are are wide difference between all of the category  and **Common mussels** was observe to be the highest.

#    SOLUTION 6
* A **pie chart** for an `attribute of your choice`

```{r}
plotdata <- planktondata %>%
  count(year) %>%
  arrange(desc(year)) %>%
  mutate(prop = round(n*100/sum(n), 1),
         lab.ypos = cumsum(prop) - 1*prop)

plotdata$label <- paste0(plotdata$year, "\n",
                         round(plotdata$prop), "%")


ggplot(plotdata, 
       aes(x = "year", 
           y = prop, 
           fill = year)) +
  geom_bar(width = 1, 
           stat = "identity", 
           color = "red") +
  geom_text(aes(y = lab.ypos, label = label), 
            color = "white", ) +
  coord_polar("y", 
              start = 0, 
              direction = 1, ) +
  theme_void() +
  theme(legend.position = "FALSE") +
  labs(title = "A pie chart for an attribute year")
```

###  COMMENT
* I observe in the graph that that some years have smaller percentage than others.
thus there is difference in **planktondata**  of each year.

#   SOLUTION 7

*  A plot to show values of **Robgordia.SP** AGAINST values of ,**Pseudonitzschia .A.SP** where the species is either `common mussels,`
or `pacific oysters.`

```{r}

planktondata %>% 
  subset(Species == "Common mussels") %>% 
  ggplot(aes(Robgordia.Sp, Pseudonitzschia.A.Sp)) +
  geom_line() +
  geom_line(data = , aes(Robgordia.Sp, Pseudonitzschia.A.Sp),color='red')+
  labs(title = "A plot to show values of Robgordia.SP AGAINST 
       values of Pseudonitzschia .A.SP",
       subtitle = "where the species is for common mussels") +
  geom_smooth(formula = y ~ x,method = "lm", se = T)
```

###     COMMENT

#    SOLUTION 8 

* A plot to show **Alexandrium.Sp** in `different` region by **farming species.**

```{r}
ggplot(planktondata, aes(Region,   Alexandrium.Sp)) + 
  geom_jitter(alpha = 0.2,color='red')+
  labs(title ="A plot to show Alexandrium.Sp in different region by farming species",
       subtitle = "Use jitter on the X- Axis")
```

###   COMMENT
* I observe that common common mussels values decrease and increase at some point. this go a long way to tell us that there are great significant difference between the two planktondata.

#    SOLUTION 9

* To determine the pair that our **Correlated and non Correlated** we first:

   + Testing the assumptions (Linearity and Normalcy) with SHAPIRo - WILK 
   
   + Correlation Coefficient
   
   + Checking for the significance
   
```{r}
###Using Shapiro test (This is a test of normality, here we are checking whether the variables are normally distributed or not )
shapiro.test(planktondata$Pseudonitzschia.A.Sp)
shapiro.test(planktondata$Alexandrium.Sp)
shapiro.test(planktondata$Robgordia.Sp)
shapiro.test(planktondata$Pseudonitzschia.A.Sp)

#Correlation Coefficient :TESTING FOR THE EFFICENCY
cor(planktondata$Pseudonitzschia.A.Sp,planktondata$Alexandrium.Sp)  
cor(planktondata$Pseudonitzschia.A.Sp,planktondata$Robgordia.Sp)   
cor(planktondata$Alexandrium.Sp,planktondata$Pseudonitzschia.A.Sp)  
cor(planktondata$Alexandrium.Sp,planktondata$Robgordia.Sp)   
cor(planktondata$Robgordia.Sp,planktondata$Pseudonitzschia.A.Sp)  
cor(planktondata$Robgordia.Sp,planktondata$Alexandrium.Sp)  

#Checking for the significance
Tes<- cor.test(planktondata$Pseudonitzschia.A.Sp,planktondata$Alexandrium.Sp,method = "pearson")
Tes
Tes0<- cor.test(planktondata$Pseudonitzschia.A.Sp,planktondata$Robgordia.Sp,method = "pearson")
Tes0
Tes1<- cor.test(planktondata$Alexandrium.Sp,planktondata$Pseudonitzschia.A.Sp,method = "pearson")
Tes1
Tes2<- cor.test(planktondata$Alexandrium.Sp,planktondata$Robgordia.Sp,method = "pearson")
Tes2
Tes3<- cor.test(planktondata$Robgordia.Sp,planktondata$Pseudonitzschia.A.Sp,method = "pearson")
Tes3
Tes4<- cor.test(planktondata$Robgordia.Sp,planktondata$Alexandrium.Sp,method = "pearson")
Tes4
```

## OR WE COULD DO ALL THREE TOGETHER
```{r}
data77 <- planktondata %>% select(Pseudonitzschia.A.Sp,Robgordia.Sp,Alexandrium.Sp)
data88 <- cor(data77,method = "pearson")
data88
summary(data88)
```

###   CONCLUSION and Justify
* Since the p-value is above 0.05 (here it is 0.09212, 2.2e-16,0.09212, and 0.07524, 
we can conclude that **Pseudonitzschia.A.Sp**, **Alexandrium.Sp** and **Robgordia.Sp** are `significantly correlated` with a value of 0.06482684,0.975273 and 0.06138349 and a p-value of 0.09212,2.2e-16,0.09212, and 0.07524.

* As we can see there is a `positive value` between **Pseudonitzschia.A.Sp**, **Alexandrium.Sp** and **Robgordia.Sp**, the point to be noted here is `correlation` is just a measure of association.

* It will tell the degree of association along with the direct or indirect proportionality.

* since all three pairing are above 0.05 , we thus conclude that the is no pair among planktondata( Pseudonitzschia.A.Sp Alexandrium.Sp Robgordia.Sp)

#   SOLUTION 10
* A `line plot` Which shows the 
**water temperature** (y axis) against the **sample index** (x axis), for *samples of common cockles and pacific oyster*.

```{r}
planktondata %>% 
  subset(Species == "Common cockles") %>% 
  ggplot(aes(Sample, Water.Temp)) +
  geom_line() +
  geom_line(data = , aes(Sample, Water.Temp),color='red')+
  labs(title = "a line plot Which shows the water temperature (y axis) against the sample index (x axis)",
       subtitle = "samples of common
cockles and pacific oyster. ") +
  geom_smooth(formula = y ~ x,method = "lm", se = T)
```

#SOLUTION 11

* `linear regression model` of **Pseudonsitzshia.A.SP** for a value of **Robgordia.sp** of 1000,2500,and 4000 *cells per liter*.
```{r}
data101 <- planktondata %>% select(Pseudonitzschia.A.Sp,Robgordia.Sp)
cor(data101,method = "pearson")
linear_regression_model <- lm(Pseudonitzschia.A.Sp ~ Robgordia.Sp,
                              data=  data101)
summary(linear_regression_model)

#linear regression model of Pseudonsitzshia.A.SP for a value of Robgordia.sp of 1000 cells per litre

cells_per_litre_1000 = -275.97689 + (9.50029*1000) 
cells_per_litre_1000

#linear regression model of Pseudonsitzshia.A.SP for a value of Robgordia.sp of 2500 cells per litre
cells_per_litre_2500 = -275.97689 + (9.50029*2500)
cells_per_litre_2500 

#linear regression model of Pseudonsitzshia.A.SP for a value of Robgordia.sp of 4000 cells per litre
cells_per_litre_4000  = -275.97689 + (9.50029*4000)
cells_per_litre_4000
```

###  justify the appropriateness of the model and comments

* We found a `significant relationship` between **Pseudonsitzshia.A.SP** and **Robgordia.sp**
(p < 2.2e-16, R2 = 0.9512), with a **Robgordia.sp** of 1000,2500,and 4000 cells per liter there is *increase*.

#   SOLUTION 12

* A `data frame with three columns`: month,year,and the mean of the **water temperatures observed in the planktondataset** during that month-year period

```{r}
planktondata=read.csv(file.choose(),header=T)
attach(planktondata)
group <- paste(month,year)
avg <- tapply(Water.Temp, group, mean)
df <- data.frame("Average temp" = avg,"Month" = as.numeric(substring(names(avg),1,2)),"Year" = as.numeric(substring(names(avg),3)),row.names = NULL)
df
```

###   Water temperature is 12 deg
* H0 -> mu = 12 (the water temperature is 12 degree)
* H1 -> mu != 12 (the water temperature is not 12 degree)

```{r}
t.test(x = Water.Temp,alternative = "two.sided",mu = 12,conf.level = .99)
```

* From the obtained P-value there is no `enough statistical evidence to reject H0` which state that the **Water temperature is 12 degrees**

* Since at 99% `level of confidence the estimated` mean of the **water temperature** is between 11.8 and 12.6 12.5, should be at the *extreme of the sampling distribution of the mean* 

* So if 12.5 should be moved to the middle of the distribution which will distort the `confidence interval` making the mean value 12.17 far off to the left of the **sampling distribution**

* Therefore when a using `type I error criteria` of 95% we'll obtain a low p-value making it easier to reject the null *hypothesis* which will come to a conclusion of the **water temperature** not being 12.5

#  SOLUTION 13

* water temperature `affected period`

* H0 -> mu1 == mu2 (Water temperature of 1st half period is equal to **water temperature** of the 2nd half period)

* H1 -> mu1 != mu2 (Water temperature of 1st half period is not equal to **water temperature** of the 2nd half period)

```{r}
t.test(Water.Temp ~ period,alternative = "two.sided",conf.level = .99)
```

* From the above analysis its `clear` that there is no difference between the **water temperature** in both the *1st  and 2nd half of the year*, so we conclude that the **water temperature** is not `affected` by the period

#   SOLUTION 14

* Our  work  is  not  done  once  we  have  found  a  `statistically  significant  difference`  between the **group means**. When we calculate $MSb$, and end up with an `average difference between the group means`. 

* Since we are comparing four group means, we might find a `relatively large average` difference between these group means even if two of the four **group means are identical**. 

* Therefore,a `statistically  significant`$F$  value  tells  us  only  that  somewhere  there  is  a `meaningful difference` between the *group means*. But it does not tell us which **groups differ** from each other `significantly`.

#    SOLUTION 15

* FOR EFFECTIVE PLOT AND NON EFFECTIVE
```{r}
plotdata <-planktondata %>%
  count(Region) %>%
  mutate(pct = n / sum(n),
         pctlabel = paste0(round(pct*100), "%"))

# plot the bars as percentages, 
# in decending order with bar labels
ggplot(plotdata, 
       aes(x = reorder(Region, -pct),
           y = pct)) + 
  geom_bar(stat = "identity", 
           fill = "indianred3", 
           color = "black") +
  geom_text(aes(label = pctlabel), 
            vjust = -0.25) +
  scale_y_continuous(labels = percent) +
  labs(x = "Region", 
       y = "Percent", 
       title  = "A bar plot for Region")

#Non effectives
plotdata <-planktondata %>%
  count(Alexandrium.Sp) %>%
  mutate(pct = n / sum(n),
         pctlabel = paste0(round(pct*100), "%"))

# plot the bars as percentages, 
# in decending order with bar labels
ggplot(plotdata, 
       aes(x = reorder(Alexandrium.Sp, -pct),
           y = pct)) + 
  geom_bar(stat = "identity", 
           fill = "indianred3", 
           color = "black") +
  geom_text(aes(label = pctlabel), 
            vjust = -0.25) +
  scale_y_continuous(labels = percent) +
  labs(x = "Alexandrium.Sp", 
       y = "Percent", 
       title  = "A bar plot for Alexandrium.Sp")

```

###  COMMENT

* the first graph can easily be read. while the second can be read but parameter and measurement are not clear.

# Rferences

* Univariate Table:  [https://www.rdocumentation.org/packages/Publish/versions/2020.12.23/topics/univariateTable](https://www.rdocumentation.org/packages/Publish/versions/2020.12.23/topics/univariateTable)

* DATA VISUALIZATION by R studio: [https://rstudio-pubs-static.s3.amazonaws.com/294957_a9f61a4e682743daa04390484e6e5a7b.html#bar_plots
](https://rstudio-pubs-static.s3.amazonaws.com/294957_a9f61a4e682743daa04390484e6e5a7b.html#bar_plots
)

* DATA VISUALIZATION WITH R 2: [https://rkabacoff.github.io/datavis/Univariate.html](https://rkabacoff.github.io/datavis/Univariate.html)

* Univariate and bivariate descriptive analysis BY:Gert Janssenswillen [https://beta.rstudioconnect.com/content/3350/dplyr_tutorial.html
](https://beta.rstudioconnect.com/content/3350/dplyr_tutorial.html
)

* [corollated https://cran.r-project.org/web/packages/psych/vignettes/intro.pdf](corollated https://cran.r-project.org/web/packages/psych/vignettes/intro.pdf)

* Stack Overflow:[https://stackoverflow.com/questions/64992231/two-ggplot-with-subset-in-pipe](https://stackoverflow.com/questions/64992231/two-ggplot-with-subset-in-pipe)